Pillar 1 - Sequential Pattern Mining

### 2019-03-12 Pillar 1 - Sequential Pattern Mining.Rmd
debug incoherence between data preparation and pillar 1
Simplify code of provide_statistics

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Pillar 1/")
```

```{r library, message = FALSE, warning = FALSE}
library(tidyverse)
library(arulesSequences)
```

# Setup
We load the environment `data_pillar1` which we saved at the end of the data preparation. It contains the data sets `d_course` and `d_transcript`.
```{r loading data}
load("Output/data_pillar_1.RDATA")
```

```{r for convenience}
pass_grade <- 5.5
high_grade <- 6.5

join_d_course <- function(data){

  data %>%
    inner_join(
    d_course,
    by = c("Course ID")
    )

join_d_transcript <- function(data){

  data %>%
    inner_join(
    d_transcript,
    by = c("Course ID")
    )

@@

# Data Exploration

We compute summary statistics (minimum, maximum, mean, median, standard deviation, failure rate, number of failure and count) at different levels (student, course, cluster, concentration, year and course level). We save the results in the environment `Transcript Statistics`.

```{r statistic summary, cache = TRUE}

#
# function providing statistics
provide_statistics <- function(data, var){

    mutate(
      Grade = case_when(
        Grade >= 5.5 ~ Grade,
        T            ~ as.numeric(NA)
        )
      ) %>%

    filter(
      !is.na(!!enquo(var))
      ) %>%
    group_by(
      !!enquo(var)
      ) %>%

    summarise(
      Count           = n(),
      Min             = min(Grade, na.rm = TRUE),
      Max             = max(Grade, na.rm = TRUE),
      Mean            = mean(Grade, na.rm = TRUE),
      Median          = median(Grade, na.rm = TRUE),
      IQR             = IQR(Grade, na.rm = TRUE),
      SD              = sd(Grade, na.rm = TRUE),
      `Low Rate`      = mean(is.na(Grade) | Grade < high_grade),
      `Fail Rate`     = mean(is.na(Grade))
      ) %>%

    mutate_at(
      vars(Mean, SD, `Low Rate`, `Fail Rate`, IQR),
      round,
      digits = 2
    )

#
# Generating Statistics
statistics <- list()

# Student level
statistics$student <- d_transcript %>%
  provide_statistics(`Student ID`)

# Course level
statistics$course <- d_transcript %>%
  join_d_course %>%
  provide_statistics(`Course ID`)

# Cluster level
statistics$cluster <- d_transcript %>%
  join_d_course %>%
  provide_statistics(Cluster)

# Type level
statistics$type <- d_transcript %>%
  join_d_course %>%
  provide_statistics(Type)

# Concentration level
statistics$concentration <- d_transcript %>%
  join_d_course %>%
  gather(
    key = X, 
    value = Concentration, 
    Concentration, `Concentration (additional)`,
    na.rm = TRUE
    ) %>%
  provide_statistics(Concentration)

# Year level
statistics$year <- d_transcript %>%
  provide_statistics(Year_numerical)

# Level level
statistics$level <- d_transcript %>%
  # TODO: filter for student who completed their studies
  join_d_course %>%
  provide_statistics(Level)
```

```{r, include = FALSE}
save(
  statistics,
  file = "Output/Transcript Statistics.RDATA"
  )
```

```{r, include = FALSE}
rm(statistics, provide_statistics)
```

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Pillar 1/")

library(tidyverse)
library(arulesSequences)
```

# Load data
We load the environment `data_pillar1` which we saved at the end of the data preparation. It contains the data sets `d_course` and `d_transcript`.
```{r loading data}
load("Output/data_general.RDATA")
load("Output/data_pillar_1.RDATA")
```

# Descriptive statistics
We compute summary statistics (minimum, maximum, mean, median, standard deviation, failure rate, number of failure and count) at different levels (student, course, cluster, concentration, year and course level). We save the results in the environment `Transcript Statistics`.

```{r statistics set up}
pass_grade <- 5.5
high_grade <- 6.5
```

```{r join_d_df}
# helper function
join_d_course     <- function(df) df %>% left_join(d_course    , by = c("Course ID"))
join_d_transcript <- function(df) df %>% left_join(d_transcript, by = c("Course ID"))
```

```{r provide_statistics}
# helper function
provide_statistics <- function(var, df = d_transcript, high_gr = high_grade, pass_gr = pass_grade, join = TRUE){

  if(join) df <- df %>% join_d_course

  df %>%

    filter(!is.na(!!enquo(var))) %>%

    group_by(!!enquo(var)) %>%

    summarise(
      Count           = n(),
      Min             = min   (Grade),
      Max             = max   (Grade),
      Mean            = mean  (Grade),
      Median          = median(Grade),
      IQR             = IQR   (Grade),
      SD              = sd    (Grade),
      `Low Rate`      = mean  (Grade < high_grade),
      `Fail Rate`     = mean  (Grade < pass_grade)
    ) %>%

    mutate_if(is.numeric, round, digits = 2)

@@

```{r statistics}
statistics <- list()

# Student level
statistics$student <- provide_statistics(`Student ID`)

# Course level
statistics$course  <- provide_statistics(`Course ID`)

# Cluster level
statistics$cluster <- provide_statistics(Cluster)

# Type level
statistics$type    <- provide_statistics(Type)

# Concentration level
statistics$concentration <- d_transcript %>%
  join_d_course %>%
  gather(
    key = X, 
    value = Concentration, 
    Concentration, `Concentration (additional)`,
    na.rm = TRUE
    ) %>%
  provide_statistics(Concentration, df = ., join = FALSE)

# Year level
statistics$year <- provide_statistics(Year_numerical)

# Level level
statistics$level <- provide_statistics(Level) # TODO: filter for student who completed their studies
```

```{r save statistics, include = FALSE}
save(statistics, file = "Output/Transcript Statistics.RDATA")
rm(statistics, provide_statistics)
```
